A DPG method for steady viscous compressible flow

### 2013-07-30 docs/Jesse/NavierStokes paper/CoverLetter.tex

%&pdflatex
\documentclass[11pt,onecolumn]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsfonts,mathrsfs,amsthm}
\usepackage[top=2cm,bottom=3cm,left=2.5cm,right=2cm]{geometry}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{array}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{graphicx}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{algorithm, algorithmic}
\usepackage{array}
\usepackage{bm}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage[normalem]{ulem}
\usepackage{epstopdf}

@@

\author{Jesse Chan\textsuperscript{a}, Norbert Heuer\textsuperscript{b}, Tan Bui-Thanh\textsuperscript{a}, and Leszek Demkowicz\textsuperscript{a}}

%\title{$\epsilon$-explicit analysis of DPG for convection-dominated diffusion}
\title{A robust DPG method for convection-dominated diffusion problems II: adjoint boundary conditions and mesh-dependent test norms.}
\date{}
\begin{document}

%\tableofcontents
\maketitle
\begin{center}
\textsuperscript{a} Institute for Computational Engineering and Sciences, \\University of Texas at Austin, \\Austin, TX 78712, USA\\
\end{center}

\begin{center}
\textsuperscript{b} Facultad de Matem\'aticas, \\Pontificia Universidad Cat\'olica de Chile,\\
Avenida Vicu\~na Mackenna 4860, Santiago, Chile
\end{center}

\section{Introduction}

\subsection{Singular perturbation problems and robustness}

The finite element/Galerkin method has been widely utilized in engineering to solve partial differential equations governing the behavior of physical phenomena in engineering problems.  The method relates the solution of a partial differential equation (PDE) to the solution of a corresponding variational problem. The finite element method itself provides several advantages --- a framework for systematic mathematical analysis of the behavior of the method, weaker regularity constraints on the solution than implied by the strong form of the equations, and applicability to very general physical domains and geometries.

Historically, the Galerkin method has been very successfully applied to a broad range of problems in solid mechanics, for which the variational problems resulting from the PDE are symmetric and coercive (positive-definite). It is well known that the finite element method produces optimal or near-optimal results for such problems, with the finite element solution matching or coming close to the best approximation of the solution in the finite element space. However, standard Bubnov-Galerkin methods tend to perform poorly for the class of PDEs known as singular perturbation problems. These problems are often characterized by a parameter that may be either very small or very large in the context of physical problems.  An additional complication of singular perturbation problems is that very often, in the limiting case of the parameter blowing up or decreasing to zero, the PDE itself will change types (e.g.\ from elliptic to hyperbolic).

\subsubsection{Convection-diffusion}

A canonical example of a singularly perturbed problem is the convection-diffusion equation.  %The equation models the convection of fluid along a direction $\beta$ with a viscous dissipation of magnitude $\epsilon$.  In the limit as $\epsilon\rightarrow 0$, the convection-diffusion equation transforms from a parabolic to a hyperbolic partial differential equation.  
In 1D, the convection-diffusion equation is
\begin{align*}
\beta u'-\epsilon u'' &= f.
\end{align*}
The equation represents the change in the concentration $u$ of a quantity in a given medium, taking into account both convective and diffusive effects. $\beta$ represents the speed of convection, while the singular perturbation parameter $\epsilon$ represents the diffusivity of the medium. In the limit of an inviscid medium as $\epsilon\rightarrow 0$, the equation changes types, from elliptic to hyperbolic, and from second order to first order. For Dirichlet boundary conditions $u(0)=u_0$ and $u(1)= u_1$, the solution can develop sharp boundary layers of width $\epsilon$ near the outflow.

%The finite element method has performed notoriously poorly for this example, as well as for many other singularly perturbed problems. This poor performance is captured by the error bound for the finite element method applied to the convection-diffusion equation - under the standard Bubnov-Galerkin method with $u\in H^1(0,1)$, we have the bound given in \cite{roos2008robust}
The poor performance of the finite element method for this problem is reflected in the bound on the error in the finite element solution --- under the standard Bubnov-Galerkin method with $u\in H^1(0,1)$, we have the bound given in \cite{roos2008robust}:
\[
\|u-u_h\|_\epsilon \leq C \inf_{w_h}\|u-w_h\|_{H^1(0,1)},
\]
for $\|u\|_\epsilon^2 \coloneqq \|u\|_{L^2}^2 + \epsilon \|u'\|_{L^2}^2$, with $C$ independent of $\epsilon$. An alternative formulation of the above bound is 
\[
\|u-u_h\|_{H^1(0,1)} \leq C(\epsilon) \inf_{w_h}\|u-w_h\|_{H^1(0,1)},
\]
where $C(\epsilon)$ grows as $\epsilon\rightarrow 0$. The dependence of the constant $C$ on $\epsilon$ is referred to as a \textit{loss of robustness} --- as the singular perturbation parameter $\epsilon$ decreases, our finite element error is bounded more and more loosely by the best approximation error.  As a consequence, the finite element solution can diverge significantly from the best finite element approximation of the solution for very small values of $\epsilon$.  For example, on a coarse mesh, and for small values of $\epsilon$, the Galerkin approximation of the solution to the convection-diffusion equation with a boundary layer develops spurious oscillations everywhere in the domain, even where the best approximation error is small.  These oscillations grow in magnitude as $\epsilon \rightarrow 0$, eventually polluting the entire solution.

%, allowing for more and more error in the finite element solution for as $\epsilon$ shrinks. 
%Intuitively, a non-robust method will perform increasingly worse as $\epsilon$ decreases.

\subsubsection{Wave propagation}

Another example of a singular perturbation problem which experiences loss of robustness is high frequency wave propagation, in which the singular perturbation parameter is the wavenumber $k$, where $k\rightarrow \infty$. The loss of robustness in this case manifests as ``pollution" error, a phenomenon in which the finite element solution degrades over many wavelengths for high wavenumbers (commonly manifesting as a phase error between the FE solution and the exact solution).

\subsubsection{Stabilization terms}

Traditionally, instability/loss of robustness has been dealt with using residual-based stabilization techniques.  Given some variational form, the problem is modified by adding to the bilinear form the strong form of the residual, weighted by a test function and scaled by a stabilization constant $\tau$.  The most well-known example of this technique is the streamline-upwind Petrov-Galerkin (SUPG) method, which is a stabilized method for solving the convection-diffusion equation using piecewise linear continuous finite elements \cite{SUPG}.  In the 1D case, SUPG stabilization not only removes the spurious oscillations from the finite element solution of the convection-diffusion equation, but delivers the best finite element approximation in the $H^1$ norm.  An important difference between residual-based stabilization techniques and other stabilizations is the idea of \textit{consistency} --- by adding stabilization terms based on the residual, the exact solution still satisfies the same variational problem (i.e.\ Galerkin orthogonality still holds).\footnote{Contrast this to an artificial diffusion method, where a specific amount of additional viscosity is added based on the magnitude of the convection and diffusion parameters. The exact solution to the original equation no longer satisfies the new stabilized formulation.}

The addition of residual-based stabilization terms has also been interpreted as a modification of the test functions --- in other words, stabilization can be achieved by changing the test space for a given problem.  We approach the idea of stabilization through the construction of \textit{optimal test functions} to achieve optimal approximation properties.

%\textcolor{red}{add bit about SUPG/residual-based stabilization terms and test functions}

\subsection{Discontinuous Petrov-Galerkin methods with optimal test functions}
\seclab{optimalTest} Petrov-Galerkin methods, in which the test space
differs from the trial space, have been explored for over 30 years,
beginning with the approximate symmetrization method of Barrett and
Morton~\cite{BARRETT01101981}. The idea was continued with the SUPG
method of Hughes, and the characteristic Petrov-Galerkin approach of
Demkowicz and Oden~\cite{Demkowicz1986188}, which introduced the
idea of tailoring the test space to change the norm in which a finite
element method would converge.

The idea of optimal test functions was introduced by Demkowicz and
Gopalakrishnan in \cite{DPG2}.  Conceptually, these optimal test
functions are the natural result of the minimization of a residual
corresponding to the operator form of a variational equation. The
connection between stabilization and least squares/minimum residual
methods has been observed previously \cite{GLS}. However, the method in \cite{DPG2} distinguishes itself
by measuring the residual of the natural \textit{operator form of the equation}, which is posed in the dual space, and measured with the dual norm, as we now discuss.

Throughout the paper, we assume that the trial space $U$ and test space $V$ are real Hilbert spaces, and denote $U'$ and $V'$ as the respective topological dual spaces. Let $U_h \subset U$ and $V_h\subset V$ be finite dimensional subsets. We are interested in the following problem  
\begin{equation}
\eqnlab{variationEq}
\left\{
  \begin{array}{l}
    \text{Given } l \in V', \text{ find } u_h \in U_h  \text{ such that} \\ 
    b(u_h,v_h) = l(v_h), \quad \forall v_h\in V_h,
  \end{array}
  \right.
\end{equation}
where $b\LRp{\cdot,\cdot}: U \times V \to \mbb{R}$ is a continuous
bilinear form.  $U_h$ is chosen to be some trial space of approximating functions, but $V_h$ is as of yet unspecified.

Throughout the paper, we suppose the variational
problem \eqnref{variationEq} to be well-posed. In that case, we can
identify a unique operator $B:U\rightarrow V'$ such that
\[
\langle Bu,v\rangle_V \coloneqq b(u,v), \quad u\in U, v\in V
\]
with $\LRa{\cdot, \cdot}_V$ denoting the duality pairing between $V'$ and $V$, to obtain the operator form of the continuous variational problem
\begin{equation}
\eqnlab{dualEq}
Bu = l \quad \text{in } V'.
\end{equation}
In other words, we can represent the continuous form of our variational equation
\eqnref{variationEq} equivalently as the operator equation \eqnref{dualEq} with values in the
dual space $V'$.  This motivates us to consider the conditions under which the solution to \eqnref{variationEq} is the solution to the minimum residual problem in $V'$ 
\[
u_h = \underset{u_h\in U_h}{\arg\min}\, J(u_h),
\]
where $J(w)$ is defined for $w\in U$ as 
\[
J(w) = \frac{1}{2}\|Bw-l\|_{V'}^2 \coloneqq\frac{1}{2} \sup_{v\in V\setminus\{0\}} \frac{| b(w,v)-l(v)|^2}{\nor{v}_V^2}.
\]
For convenience in writing, we will abuse the notation $\sup_{v \in V}$ to denote $\sup_{v\in V\setminus\{0\}}$ for the remainder of the paper.

Let us define $R_V: V \to V'$ as the Riesz map, which identifies
elements of $V$ with elements of $V'$ by 
\[
\langle R_V v,\delta
v\rangle_V \coloneqq(v, \delta v)_V, \quad \forall \delta v \in V.
\]
Here, $(\cdot, \cdot)_V$ denotes the
inner product in $V$. As $R_V$ and its inverse, $R_V^{-1}$, are both
isometries, e.g.\ $\|f\|_{V'} = \|R_V^{-1} f\|_V, \forall f \in V'$, we
have
\begin{equation}
\eqnlab{minimization}
\min_{u_h\in U_h} J(u_h) = \frac{1}{2}\left\|Bu_h-l\right\|_{V'}^2 =  \frac{1}{2}\left\|R_V^{-1}(Bu_h-l)\right\|_V^2.
\end{equation}
The first order optimality condition for \eqnref{minimization} requires
the G\^ateaux derivative to be zero in all directions $\delta u \in
U_h$, i\.e\.,
\begin{align*}
\left(R_V^{-1}(Bu_h-l),R_V^{-1}B\delta u\right)_V = 0, \quad \forall \delta u \in U. 
\end{align*}
We define, for a given $\delta u \in U$, the corresponding {\em optimal test function} $v_{\delta u}$
\begin{equation}
\eqnlab{optv}
v_{\delta u} \coloneqq R_V^{-1}B\delta u \quad  \text{in } V.
\end{equation} 
The optimality condition then becomes
\[
 \langle Bu_h-l, v_{\delta u}\rangle_V = 0, \quad \forall \delta u \in U
\]
which is exactly the standard variational equation in
 \eqnref{variationEq} with $v_{\delta u}$ as the test functions. We can define the optimal test space $V_{\rm opt} \coloneqq \{v_{\delta u} \text{ s.t. } \delta u\in U\}$. Thus, the solution of the variational problem \eqnref{variationEq} with test space $V_h = V_{\rm opt}$ minimizes the residual in the dual norm $\nor{Bu_h-l}_{V'}$. This is the key idea behind the concept of optimal test functions.

Since $U_h \subset U$ is spanned by a finite number of basis functions $\LRc{\varphi_i}_{i=1}^N$, \eqnref{optv} allows us to compute (for each basis function) a corresponding optimal test function $v_{\varphi_i}$. The collection $\LRc{v_{\varphi_i}}_{i = 1}^N$ of optimal test functions then forms a basis for the optimal test space.  In order to express optimal test functions defined in \eqnref{optv} in a more familiar form, we take  $\delta u = \varphi$, a generic basis function in $U_h$, and rewrite \eqnref{optv} as
\[
R_Vv_{\varphi} = B\varphi, \quad \text{in } V',
\]
which is, by definition, equivalent to
\[
\LRp{v_\varphi,\delta v}_V = \LRa{R_Vv_\varphi,\delta v}_{V}=
\LRa{B\varphi, \delta v}_V = b\LRp{\varphi,\delta v}, \quad
\forall \delta v \in V.
\]
%% Now, let us define the \textit{trial-to-test} operator $T =
%% R_V^{-1}B$, which, by \eqnref{optv}, maps a trial function $\delta u$
%% to its corresponding \textit{optimal} test function $v = T\delta u$.
%% On the other hand,
As a result, optimal test functions can be determined by solving the auxiliary
variational problem
\begin{equation}
\eqnlab{optvVar}
\left(v_\varphi,\delta v\right)_V = b(\varphi,\delta v), \quad \forall
\delta v \in V.
\end{equation}
However, in general, for standard $H^1$ and $H({\rm div})$-conforming finite element methods, test functions are continuous over the entire domain, and hence solving variational problem \eqnref{optvVar} for each optimal test function requires a global operation over the entire mesh, rendering the method impractical. A breakthrough came through the development of discontinous Galerkin (DG) methods, for which basis functions are discontinuous over elements. In particular, the use of discontinuous test functions $\delta v$ and a \textit{localizable} norm $\nor{\cdot}_V$\footnote{A localizable norm $\nor{v}_{V(\Oh)}$ can be written in the form 
$$\nor{v}_{V(\Oh)}^2 = \sum_{K\in\Oh} \nor{v}_{V(K)}^2,$$ where $\nor{v}_{V(K)}$ is a norm over the element $K$.} reduces the problem of determining global optimal test functions in \eqnref{optvVar} to local problems that can be solved in an element-by-element fashion.

We note that solving \eqnref{optvVar} on each element exactly is infeasible since it amounts to inverting the Riesz map $R_V$ exactly. Instead, optimal test functions are approximated using the standard Bubnov-Galerkin method on an ``enriched" subspace $\tilde{V} \subset V$ such that $\dim(\tilde{V}) > \dim(U_h)$ elementwise \cite{DPG1, DPG2}. In this paper, we assume the error in approximating the optimal test functions is negligible, and refer to the work in \cite{practicalDPG} for estimating the effects of approximation error on the performance of DPG.

It is now well known that the DPG method delivers the best approximation error in the ``energy norm" --- that is \cite{Bui-ThanhDemkowiczGhattas11a, DPG2,DPG4} 
\begin{equation}
\eqnlab{optimalError}
\nor{u-u_h}_{U,E} = \inf_{w\in U_h} \nor{u-w}_{U,E},
\end{equation}
where the energy norm $\|\cdot \|_{U,E}$ is defined for a function $\varphi \in U$ as
\begin{equation}
\eqnlab{energyNorm} \|\varphi\|_{U,E} \coloneqq \sup_{v\in V}
\frac{b(\varphi,v)}{\|v\|_V} = \sup_{\nor{v}_V = 1} b(\varphi,v) =
\sup_{\nor{v}_V = 1} \LRa{B\varphi,v}_V = \nor{B\varphi}_{V'} =
\nor{v_\varphi}_V,
\end{equation}
where the last equality holds due to the isometry of the Riesz map
$R_V$ (or directly from \eqnref{optvVar} by taking the supremum). An
additional consequence of adopting such an energy norm is that,
without knowing the exact solution, the energy error $\|u-u_h\|_{U,E}  = \nor{Bu - Bu_h}_{V'} = \nor{R_V^{-1}\LRp{l-Bu_h}}_{V}$ can
be determined by computing the \textit{error representation function} $e \coloneqq R_V^{-1}\LRp{l-Bu_h}$ through
\[
\left(e,\delta v\right)_V = b(u-u_h,\delta v) = l\LRp{\delta v} - b(u_h,\delta v)
\]
and measuring its norm $\|e\|_V$.  This is simply a consequence of the least-squares nature of DPG; the energy error is simply the norm of the residual in $V'$.  Under the assumption of a localizable norm on $V$, we can compute the squared norm over the entire domain $\nor{e}_{V(\Oh)}^2$ as the sum of individual element contributions $\sum_{K\in \Oh} \nor{e}_{V(K)}^2$.  We define $e_K^2 \coloneqq \nor{e}_{V(K)}^2$ as a local error indicator with which we can drive adaptive mesh refinement.

Practically speaking, this implies that the DPG method is discretely stable on any mesh. In particular, DPG is unconditionally stable for higher order adaptive meshes, where discrete stability is often an issue.

\subsection{Duality between trial and test norms (energy norm pairings)}
%\subsection{Abstract energy norm pairings}
\seclab{energyPair}
A clear property of the energy norm defined by \eqnref{energyNorm} is that the trial norm $\nor{\cdot}_{U,E}$ is induced by a given test norm. However, the reverse relationship holds as well; for any trial norm, the test norm that induces such a norm is recoverable through duality. We have a result, Lemma 2.5 in \cite{Bui-ThanhDemkowiczGhattas11a}: assuming, for simplicity, that the bilinear form $b(u,v)$ is definite\footnote{By definite, we mean that \begin{align*}b(u,v) &= 0, \forall v\in V \Rightarrow u = 0\\ b(u,v)&=0, \forall u\in U \Rightarrow v = 0,\end{align*}  which imply injectivity of the bilinear operator $B$ and its transpose $B'$, defined such that $\LRa{Bu,v} = \LRa{u,B'v}$.  These conditions imply solvability of the variational problem.}, given any norm $\nor{\cdot}_{U}$ on the trial space $U$, for $\varphi \in U$, we can represent $\nor{\varphi}_{U}$ via
\[
\nor{\varphi}_{U} = \sup_{v \in V}\frac{b\LRp{w,v}}{\nor{v}_{V,U}}.
\]
where $\nor{v}_{V,U}$ is defined through
\[
\nor{v}_{V,U} = \sup_{w \in U}\frac{b\LRp{w,v}}{\nor{w}_{U}}.
\]
%In other words, the norm $\nor{\cdot}_V$ in $V$ can be recovered using the energy norm $\nor{\cdot}_{U,E}$ in $U$, and vice versa.

In particular, given two arbitrary norms $\nor{\cdot}_{U,1}$ and $\nor{\cdot}_{U,2}$ in $U$
such that $\nor{\cdot}_{U,1} \le c \nor{\cdot}_{U,2}$ for some constant
$c$, they generate two norms $\nor{\cdot}_{V,U,1}$ and
$\nor{\cdot}_{V,U,2}$ in $V$ defined by
\[
\nor{v}_{V,U,1} \coloneqq \sup_{w \in U}\frac{b\LRp{w,v}}{\nor{w}_{U,1}}, \quad
\text{and }\nor{v}_{V,U,2} \coloneqq \sup_{w \in U}\frac{b\LRp{w,v}}{\nor{w}_{U,2}},
\]
such that $\nor{\cdot}_{V,U,1}$ and $\nor{\cdot}_{V,U,2}$ induce
$\nor{\cdot}_{U,1}$ and $\nor{\cdot}_{U,2}$ as energy
norms in $U$, respectively. That is,
\[
\|\varphi\|_{U,1} = \sup_{v\in V}
\frac{b(\varphi,v)}{\|v\|_{V,U,1}}, \quad \text{and }\|\varphi\|_{U,2} = \sup_{v\in V}
\frac{b(\varphi,v)}{\|v\|_{V,U,2}}.
\]

A question that remains to be addressed is to establish the relationship
between $\nor{\cdot}_{V,U,1}$ and $\nor{\cdot}_{V,U,2}$, given that
$\nor{\cdot}_{U,1} \le c \nor{\cdot}_{U,2}$. But this is
straightforward since we have
\begin{align*}
 \| v \|_{V,U,2} = \sup_{u \in U} \frac{b\left(w,v\right)}{\left\|
  w \right\|_{U,2}} \le c\sup_{w \in U} \frac{b\left(w,v\right)}{\left\| w
  \right\|_{U,1}} = c\| v \|_{V,U,1}.
\end{align*}
Consequently, a stronger energy norm in $U$ will generate a weaker
norm in $V$ and vice versa. In other words, to show that an
energy norm $\nor{\cdot}_{U,1}$ is weaker than another energy norm
$\nor{\cdot}_{U,2}$ in $U$, one simply needs to show the reverse inequality on the
corresponding norms in $V$, that is, $\nor{\cdot}_{V,U,1}$ is stronger
than $\nor{\cdot}_{V,U,2}$.

%From now on, unless otherwise stated, we will refer to  $\nor{\cdot}_{V,U,1}$, $\nor{\cdot}_{V,U,2}$, and $\nor{\cdot}_{V}$ as the test norms which induce $\nor{\cdot}_{U,1}$, $\nor{\cdot}_{U,2}$, and $\nor{\cdot}_{U,E}$ and vice versa, in the sense discussed above. 
From now on, unless otherwise stated, we will refer to $\nor{\cdot}_{V,U}$ as the test norm that induces a given norm $\nor{\cdot}_U$. Likewise, we will refer $\nor{\cdot}_{U,V}$ as the trial norm induced by a given test norm $\nor{\cdot}_V$. In this paper, for simplicity of exposition, we shall call a pair of norms in $U$ and $V$ that induce each other as an {\em energy norm pairing}.

\subsection{Discontinuous Petrov-Galerkin methods with the ultra-weak formulation}
\seclab{abstractUweak}

The naming of the discontinuous Petrov-Galerkin method refers to the fact that the method is a Petrov-Galerkin method, and that the test functions are specified to be discontinuous across element boundaries. There is no specification of the regularity of the trial space, and we stress that the idea of DPG is not inherently tied to a single variational formulation \cite{Bui-ThanhDemkowiczGhattas11a}. Additionally, Cohen, Dahmen and Welper simultaneously extend the minimum residual concept behind DPG to general variational settings and avoid the use of discontinuous test functions by formulating the minimum residual method as a saddle-point problem \cite{DahmenVariationalStabilization}.

In most of the DPG literature, however, the discontinuous Petrov-Galerkin method refers to the combination of the concept of locally computable optimal test functions in Section \secref{optimalTest} with the so-called ``ultra-weak formulation" \cite{DPG1,DPG2,DPG3,DPG4,DPGElas,DBLP:journals/procedia/NiemiCC11}. Unlike the previous two sections in which we studied the general equation \eqnref{variationEq} given by abstract bilinear and linear forms, we now consider a concrete instance of \eqnref{variationEq} resulting from an ultra-weak formulation for an abstract first-order system of PDEs $Au = f$. Additionally, from this section onwards, we will refer to DPG as the pairing of the ultra-weak variational formulation with the concept of locally computable optimal test functions.

We begin by partitioning the domain of interest $\Omega$ into $\Nel$ non-overlapping elements $K_j, j = 1,\hdots,\Nel$ such that $\Oh = \cup_{j=1}^\Nel K_j$ and $\overline{\Omega} = \overline{\Omega}_h$. Here, $h$ is defined as $h= \max_{j\in \LRc{1,\hdots,\Nel}}\text{diam}\LRp{K_j}$.  We denote the mesh ``skeleton" by $\Gh = \cup_{j=1}^\Nel \partial K_j$; the set of all faces/edges $e$, each of which comes with a normal vector ${n}_e$. The internal skeleton is then defined as $\Gamma^0_h = \Gh \setminus \partial \Omega$. If a face/edge $e \in \Gh$ is the intersection of $\partial K_i$ and $\partial K_j$, $i \ne j$, we define the following jumps:
\[
\jump{v} = \text{sgn} \LRp{{n}^-}v^- + \text{sgn} \LRp{{n}^+}v^+, \quad
\jump{\tau \cdot n} = {n}^-\cdot \tau^- + {n}^+\cdot\tau^+,
\]
where
\[
\text{sgn}\LRp{{n}^{\pm}} =
\left\{
\begin{array}{ll}
1 & \text{if } {n}^{\pm} = {n}_e \\
-1 & \text{if } {n}^\pm = -{n}_e
\end{array}
\right..
\]
For $e$ belonging to the domain boundary $\partial \Omega$, we define
\[
\jump{v} = v, \quad
\jump{\tau \cdot n} = {n}_e\cdot \tau.
\]
Note that we allow arbitrariness in assigning ``-'' and ``+'' quantities to the adjacent elements $K_i$ and $K_j$.
%For the rest of the paper, we will use the same notation for both a function and its trace (if it is well-defined) when there is no ambiguity.

The ultra-weak formulation for $Au = f$ on $\Oh$, ignoring boundary
conditions for now, reads
\begin{equation}
\eqnlab{uweak}
b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \langle \widehat{u}, \jump{v}
\rangle_{\Gh} - (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh},
\end{equation}
where we have denoted $\LRa{\cdot,\cdot}_{\Gh}$ as the duality
pairing on $\Gh$, $\LRp{\cdot,\cdot}_{\Oh}$ the $L^2$-inner
product over $\Oh$, and $A_h^*$ the formal adjoint resulting from
element-wise integration by parts.  Occasionally, for simplicity in
writing, we will ignore the subscripts in the duality pairing and
$L^2$-inner product if they are $\Gh$ and $\Oh$. Both the
inner product and formal adjoint are understood to be taken
element-wise. Using the ultra-weak formulation, the regularity
requirement on solution variable $u$ is relaxed, that is, $u$ is now
square integrable for the ultra-weak formulation \eqnref{uweak} to be
meaningful, instead of being (weakly) differentiable.  The trade-off
is that $u$ does not admit a trace on $\Gh$ even though it did
originally. Consequently, we need to introduce an additional new
``trace'' variable $\widehat{u}$ in \eqnref{uweak}, that is defined only on
$\Gh$.

The energy setting is now clear; namely,
\[
u\in L^2\LRp{\Oh} \equiv L^2(\Omega), \quad v\in V=D(A^*_h), \quad
\widehat{u}\in \gamma(D(A)),
\]
where $D(A_h^*)$ denotes the broken graph space corresponding to $A_h^*$,
and $\gamma(D(A))$ the trace space (assumed to exist) of the graph space of
operator $A$. The first discussion of the well-posedness of DPG with the ultra-weak formulation can be found in \cite{analysisDPG}, where the proof is presented for the Poisson and convection-diffusion equations. A more comprehensive discussion of the abstract setting for DPG with the ultra-weak formulation using the graph space, as well as a more general proof of well-posedness, can be consulted in \cite{Bui-ThanhDemkowiczGhattas11b}.

%\subsection{Optimal and quasi-optimal test norms}
\subsection{A canonical energy norm pairing for ultra-weak formulation}

From the discussion in Section~\secref{energyPair} of energy norm and test norm pairings, we know that specifying either a test norm or trial norm is sufficient to define an energy pairing. %In this section, we derive and discuss two important energy norm pairings, the first of which begins by specifying the canonical norm in $U$ and inducing a test norm on $V$. The second pairing begins instead by specifying the canonical norm on $V$ under the ultra-weak formulation \eqnref{uweak} and inducing an energy norm on the trial space $U$.
In this section, we derive and discuss an important energy norm pairing which specifies the canonical norm in $U$ and induces a test norm on $V$. %The second pairing begins instead by specifying the canonical norm on $V$ under the ultra-weak formulation \eqnref{uweak} and inducing an energy norm on the trial space $U$.

We begin first with the canonical norm in $U$. Since $\uh \in \gamma\LRp{D\LRp{A}}$, the standard norm for $\uh$ is
the so-called minimum energy extension norm defined as
\begin{equation}
\eqnlab{MEnorm}
\|\widehat{u}\| = \inf_{w\in D\LRp{A},
  \left.w\right|_{\Gh}=\widehat{u}} \|w\|_{D\LRp{A}}.
\end{equation}
The canonical norm for the group variable $\LRp{u,\uh}$ is then given by
\[
\|\left(u,\widehat{u}\right)\|_U^2 = \|u\|^2_{\L} + \|\widehat{u}\|^2.
\]
Applying the Cauchy-Schwarz inequality, we arrive at
\[
b\LRp{\LRp{u,\uh},v} \le \nor{\LRp{u,\uh}}_U \nor{v}_{V,U},
\]
where
\[
\nor{v}_{V,U}^2 = \|A_h^*v\|_{\L}^2
+\left(\sup_{\widehat{u} \in \gamma\LRp{D\LRp{A}}} \frac{\LRa{ \widehat{u},
  \jump{v} }_{\Gh}}{\|\widehat{u}\|}\right)^2.
\]

%On the other hand, since $v \in D\LRp{A^*_h}$, the canonical norm for
%$v$ is the broken graph norm: 
%\[
%\nor{v}_V^2 =  \|A_h^*v\|_{\L}^2 + \nor{v}_{\L}^2.
%\]
%Using the Cauchy-Schwarz inequality again, we obtain
%\[
%b\LRp{\LRp{u,\uh},v} \le \nor{\LRp{u,\uh}}_{U,V} \nor{v}_{V},
%\]
%where
%\begin{equation}
%\eqnlab{inducedQuasi}
%\nor{\LRp{u,\uh}}_{U,V}^2 = \nor{u}_{\L}^2
%+\sup_{v \in D\LRp{A_h^*}} \frac{\LRa{\widehat{u},
%  \jump{v}}_{\Gh}^2}{\|v\|_V^2},
%\end{equation}

Using the framework developed in \cite{Bui-ThanhDemkowiczGhattas11a}, one can show that $\LRp{\nor{\LRp{u,\uh}}_U,  \nor{v}_{V,U}}$ is an energy norm pairing in the sense discussed in Section \secref{energyPair}. That is, the canonical norm $\nor{\LRp{u,\uh}}_U$ in $U$ induces (generates) the norm $\nor{v}_{V,U}$ in $V$.

The canonical norm $\nor{\LRp{u,\uh}}_U$ in $U$ provides an optimal balance between the standard norms on the field $u$ and the flux $\uh$ \cite{DPG4}. As a result, if the induced norm $\nor{v}_{V,U}$ (namely, the optimal test norm) is used to compute  optimal test functions in \eqnref{optvVar}, the finite element error in the canonical norm is the best in the sense of \eqnref{optimalError}. Unfortunately, the optimal test norm is non-localizable due to the presence of the jump term $\jump{v}$. Since the jump terms couple elements together, the evaluation of the jump terms requires contributions from all the elements in the mesh. Consequently, solving for an optimal test function amounts to inverting the Riesz map over the entire mesh $\Oh$, making the optimal test norm impractical.

On the other hand, since $v \in D\LRp{A^*_h}$, we can use as a test norm for $v$ the broken graph norm: 
\[
\nor{v}_V^2 =  \|A_h^*v\|_{\L}^2 + \nor{v}_{\L}^2.
\]
This norm is a localization of $\nor{v}_{V,U}$ to allow for the solution of optimal test functions on an element-by-element basis, and is considered to be the canonical norm on $V$.  In the DPG literature \cite{DPG4}, $\nor{v}_{V,U}$ is known as the {\em optimal test norm}, while $\nor{v}_{V}$ is known as the {\em quasi-optimal} or {\em graph test norm}.

%Using the framework developed in \cite{Bui-ThanhDemkowiczGhattas11a}, one can show that both pairs $\LRp{\nor{\LRp{u,\uh}}_U,  \nor{v}_{V,U}}$ and $\LRp{\nor{\LRp{u,\uh}}_{U,V}, \nor{v}_{V}}$ are energy norm pairings in the sense discussed in Section \secref{energyPair}. That is, the canonical norm $\nor{\LRp{u,\uh}}_U$ in $U$ induces (generates) the norm $\nor{v}_{V,U}$ in $V$, while the canonical norm $\nor{v}_V$ in $V$ induces (generates) the energy norm $\nor{\LRp{u,\uh}}_{U,V}$ in $U$. In the DPG literature \cite{DPG4}, $\nor{v}_{V,U}$ is known as the {\em optimal test norm}, while $\nor{v}_{V}$ is known as the {\em quasi-optimal test norm}.

>>>>

%\documentclass[11pt,onecolumn]{scrartcl}
\documentclass{letter}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsfonts,mathrsfs,amsthm}
\usepackage[top=2cm,bottom=3cm,left=2.5cm,right=2cm]{geometry}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{array}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{graphicx}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{algorithm, algorithmic}
\usepackage{array}
\usepackage{bm}
\usepackage{tikz}
%\usepackage{subfigure}
\usepackage[normalem]{ulem}

@@

\signature{Jesse Chan, John A. Evans}
\address{201 East 24th St, Stop C0200\\
Austin, Texas 78712-1229}
\begin{document}
\begin{letter}{Drs  Chunlei Liang, Krzysztof Fidkowski, Per-Olof Persson, and Peter Vincent\\Computers and Fluids Co-guest editors\\Special Issue:12th USNCCM}

%\tableofcontents
%\maketitle

\opening{Dear Dr. Liang, Dr. Fidkowski, Dr. Persson, and Dr Vincent.,}

Please find enclosed our manuscript, 
\begin{center}
Jesse Chan, Leszek Demkowicz, Robert Moser\\
\textit{A DPG method for steady viscous compressible flow},
\end{center}
which we would like to submit for publication as an original research article in the 12th USNCCM special issue of Computers and Fluids.  We believe this is the first report describing the application of the Discontinuous Petrov-Galerkin method of Demkowicz and Gopalakrishnan to the higher order adaptive solution of problems with shocks and boundary layers in viscous compressible flow.  The method and results of this paper have been accepted for presentation at the minisymposium on \textit{Higher order methods for CFD} at the 12th USNCCM conference, and we believe that the method and results discussed in this manuscript would appeal to the readership of Computers and Fluids.

All authors have approved the manuscript and agree with its submission, and we confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.  We look forward to hearing from you at your earliest convenience.

\closing{Best regards}

\end{letter}
\end{document}
